Intrusion Detection with \\ Genetic Algorithms and Fuzzy Logic

### 2013-11-06 Talk/Presentation.tex
Added titles to each slide, also wrote things on the RLD09 slide and the rules slide, and started the fuzzy logic slide.

\subsection*{The big picture}

\begin{frame}
  \frametitle{The big picture}

@@

\begin{frame}
  \frametitle{Outline}
  \tableofcontents[hideallsubsections]
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Background]{Background}
\subsection{Types of Networking Attacks}
\begin{frame}
  \frametitle{title here}

\subsection{Detection Methodologies}
\begin{frame}
  \frametitle{title}

@@

\begin{frame}
  \frametitle{Some features of KDD99}
	\begin{enumerate}
		\item duration: length of the normal or attack activity
in seconds.
		\item src\_bytes: number of bytes sent from source to destination.
		\item num\_failed\_logins: number of failed login attempts.
		\item root\_shell: returns 1 if root shell is obtained, else returns 0.
		\item num\_access\_files: number of operations on access control files.
		\item srv\_count: number of connections to the same service as the current connection in the past two seconds.
		\item serror\_rate: percentage of connections that have ``SYN" errors.
		\item same\_srv\_rate: percentage of connections to the same service.
	\end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{RLD09}

@@

\subsection{Rules}
\begin{frame}
  \frametitle{title}

@@

\subsection{Fuzzy Logic}
\begin{frame}
  \frametitle{title}

@@

\subsection{Genetic Algorithm}
\begin{frame}
  \frametitle{title}

\subsection{Determining the accuracy of an algorithm}
\begin{frame}
  \frametitle{title}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Genetic Algorithm Implementation]{Genetic Algorithm Implementation}
\subsection{Algorithm Overview}
\begin{frame}
  \frametitle{title here}

\subsection{Experimental Design and Results}
\begin{frame}
  \frametitle{title}

@@

@@

\begin{frame}
  \frametitle{Measuring the probability of a record being an attack}

@@

\begin{frame}
	\frametitle{Encoding of features and rules}
	\begin{itemize}
	\item The four parameters are encoded into blocks.
	\item Each block is a feature with values between 0.0 and 7.0.
	\end{itemize}

@@

\begin{frame}
	\frametitle{title}

@@

\begin{frame}
	\frametitle{title}

>>>>

\subsection*{The Big Picture}

\begin{frame}
  \frametitle{The Big Picture}

@@

\begin{frame}
  \frametitle{Outline}
  \tableofcontents[hideallsubsections]
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Background]{Background}
\subsection{Types of Networking Attacks}
\begin{frame}
  \frametitle{Types of Networking Attacks}
Explain DoS, remote to user, user to root, probe
\end{frame}

\subsection{Detection Methodologies}
\begin{frame}
  \frametitle{Detection Methodologies}
Explain signature-based and anomaly-based detection
\end{frame}

@@

\begin{frame}
  \frametitle{Some Features of KDD99}
	\begin{enumerate}
		\item duration: length of the normal or attack activity
in seconds.
		\item src\_bytes: number of bytes sent from source to destination.
		\item num\_failed\_logins: number of failed login attempts.
		\item root\_shell: returns 1 if root shell is obtained, else returns 0.
		\item num\_access\_files: number of operations on access control files.
		\item srv\_count: number of connections to the same service as the current connection in the past two seconds.
		\item serror\_rate: percentage of connections that have ``SYN" errors.
		\item same\_srv\_rate: percentage of connections to the same service.
	\end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{RLD09}
	\begin{itemize}
		\item RLD09 was created because KDD99 is 14 years old.
		\item Data was captured from a university in Bangkok, Thailand.
		\item The data has 10 million data packets.
		\item 17 different types of attacks - divided into denial of service attacks, probe attacks. It also has normal activity.
		\item 12 features, which include the number of packets, source ports, and destination ports.
	\end{itemize}
\end{frame}

\subsection{Rules}
\begin{frame}
  \frametitle{Rules}
	\begin{itemize}
		\item Elements of one set are separated into different sets in order to differentiate between normal connections and attacks.

		\item If <\emph{condition}> then <\emph{action}>.

		\item Specify the details of a packet such as the IP address or port number.
		\item If a packet matches any of the rules in the intrusion detection system, the system will take appropriate action, which may include stopping the connection or logging off the system.

\subsection{Fuzzy Logic}
\begin{frame}
  \frametitle{Fuzzy Logic}
  \begin{columns}
  \begin{column}{0.6\textwidth}
  \begin{itemize}
  	\item Used to detect patterns that have a behavior that is between normal and unusual.
  	\item Fuzzy logic rule: If <\emph{condition}> then <\emph{consequence}>.
  	\begin{itemize}
  		\item \emph{condition} is a fuzzy variable
  		\item \emph{consequence} is a fuzzy set
  	\end{itemize}
  \end{itemize}  
  \end{column}

  \begin{column}{0.4\textwidth}
  \includegraphics[width=0.95\textwidth]{../triangleFigure.JPG}
  \end{column}
  \end{columns}
\end{frame}

\subsection{Genetic Algorithms}
\begin{frame}
  \frametitle{Genetic Algorithms}

@@

\subsection{Determining the Accuracy of an Algorithm}
\begin{frame}
  \frametitle{Determining the Accuracy of an Algorithm}
Explain training and test set, false positive, false negative, true positive, true negative, detection rate.
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Genetic Algorithm Implementation]{Genetic Algorithm Implementation}
\subsection{Algorithm Overview}
\begin{frame}
  \frametitle{Algorithm Overview}

@@

\subsection{Experimental Design and Results}
\begin{frame}
  \frametitle{Experimental Design}

@@

\begin{frame}
  \frametitle{Results}

@@

\subsection{Main Points of Research}
\begin{frame}
  \frametitle{Main Points of Research}
	\begin{itemize}
		\item Detecting new or unknown types of attacks in a network.
		\item The intrusion detection system used is able to identify normal network activity as well as attacks using a fuzzy genetic algorithm.

		\item Ran experiments using only RLD09, and experiments using KDD99 and RLD09 together.
	\end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Measuring the Probability of a Record Being an Attack}

@@

\begin{frame}
	\frametitle{Encoding of Features and Rules}
	\begin{itemize}
	\item The four parameters are encoded into blocks.
	\item Each block is a feature with values between 0.0 and 7.0.
	\end{itemize}

@@

\begin{frame}
	\frametitle{Algorithm Overview}

@@

\begin{frame}
	\frametitle{Experiments Using Only RLD09}
Experiment 1

\begin{frame}
	\frametitle{Experiments Using Only RLD09}
Experiment 2

\begin{frame}
	\frametitle{Experiments Using Both RLD09 and KDD99}
Experiment 1

\begin{frame}
	\frametitle{Experiments Using Both RLD09 and KDD99}
Experiment 2

\begin{frame}
	\frametitle{Experiments Using Both RLD09 and KDD99}
Experiment 3
